---
title: "Practical Machine Learning Project"
author: "YS"
date: "Monday, February 16, 2015"
output: html_document
---

## Executive Summary
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

In this project, my goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants to predict the manner in which they did the exercise.

## Getting and Cleaning Data 
First, load `pml-training` and `pml-testing` datasets, which each contains 160 variables.Perform data cleanup using the following steps: 

1. Delete the columns that contain `NA`.
2. Delete columns #1-7, as they only contain information about user name, timestamps, and windows.

```{r echo=TRUE} 
  library(randomForest)
  library(caret)
  setwd("C:/Users/Yuewei/Documents/R/data/Practical_Machine_Learning")
  trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
  testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
  training_data <- read.csv(url(trainUrl), header=TRUE, na.string=c("NA","#DIV/0!",""))
  evaluation_data <- read.csv(url(testUrl), header=TRUE, na.string=c("NA","#DIV/0!",""))
  training_clean <- training_data[, colSums(is.na(training_data))==0]
  training_clean <- training_clean[-c(1:7)]
  evaluation_clean <- evaluation_data[, colSums(is.na(evaluation_data))==0]
  evaluation_clean <- evaluation_clean[-c(1:7)]
```

##Build A Random Forest Model
First, create training and testing sets from the clean training data, in order to perform cross validation. 

```{r echo=TRUE} 
  inTrain <- createDataPartition(y=training_clean$classe, p=0.75, list=FALSE)
  training <- training_clean[inTrain,]
  testing <- training_clean[-inTrain,]
```

Second, build a random forest using `randomForest` package. 

```{r echo=TRUE}
  modFit1 <- randomForest(classe ~., data=training)
  modFit1
```

##Perform Cross Validation
I then perform cross validation by comparing the predicted classes of the testing set to the actual classes. 

```{r echo=TRUE}
  pred1 <- predict(modFit1, testing) 
  confusionMatrix(pred1, testing$classe)
```

As shown above, the `accuracy` of this model is very high (99.61%). Therefore, I can use it to predict the classes for the test cases. 

##Predict Test Cases

Here, the random forest model built above was used to predict the 20 different test cases. As the in-sample error is 0.39% (`1 - accuracy`), the out-of-sample error is estimated to be 0.39%.

```{r echo=TRUE}
  pred2 <- predict(modFit1, evaluation_clean)

  pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

  pml_write_files(pred2)
```

Conclusion: The model sucessfully predicted 20 out of 20 test cases. 